---
title: "IMDB_network_final"
output: pdf_document
---

```{r include = F}
library(GGally)
library(dplyr)
library(ggplot2)
library(stringr)
library(plyr)
library(statnet)
library(RColorBrewer)
```

```{r}
devtools::install_github("briatte/ggnet")
library(ggnet)
library(network)
library(sna)
```


#Cleaning the Data
```{r}
imdb = read.csv("imdb.csv", header = T, na.strings=c("","NA"))
awards = read.csv("awards.csv", header = T, na.strings = c("", "NA"))
cpi = read.csv("cpi.csv", header = T)
colnames(awards) = tolower(colnames(awards))

names(cpi) = c("year", "cpi")
colSums(is.na(imdb))
```

Clearly there are a significant number of missing observations in this data set.  Because this analysis is focused only on creating a network plot, we can omit these rows with NAs.

```{r}

imdb = na.omit(imdb)
imdb = imdb %>% select(gross, genres, movie_title, country, movie_imdb_link, budget, title_year, imdb_score, content_rating )

names(imdb) = c("gross", "genres", "title", "country","links", "budget", "year", "score", "rating")

#simplify the genres by taking the first entry
imdb$genres = as.character(imdb$genres)
imdb$genres_simple = strsplit(imdb$genres, split = "|", fixed = TRUE) 
imdb$genres_simple = as.character(imdb$genres_simple)
imdb$genres_simple = str_extract(imdb$genres, pattern = "^[A-Za-z]{1,20}")
imdb$genres_simple = as.factor(imdb$genres_simple)
imdb$title = gsub(imdb$title, pattern = "?", replacement ="")
imdb$links = as.character(imdb$links)

#Convert data types
imdb$genres = as.factor(imdb$genres)
imdb$budget = as.numeric(imdb$budget)
imdb$gross = as.numeric(imdb$gross)
imdb$score = as.numeric(imdb$score)
imdb$rating = as.factor(imdb$rating)

levels(imdb$genres_simple)
link_pat = '(.){35}'
imdb$links = str_extract(imdb$links, pattern = link_pat)

```

#Using CPI data to convert to adjust for inflation

This uses data from the US census to adjust for inflation.  I also filter out movies not made in the USA.

```{r}
imdb = inner_join(imdb, cpi, by = "year")
reference_year_cpi = filter(imdb, year ==2016)$cpi[1]
imdb$gross_adj = reference_year_cpi / imdb$cpi

imdb= imdb %>%
  group_by(year) %>%
  mutate(cpi_ratio=  reference_year_cpi/cpi) %>%
  mutate(gross_adj = gross*cpi_ratio) %>%
  mutate(budget_adj = budget*cpi_ratio) %>%
  select(-gross, -budget) %>%
  ungroup() %>%
  filter( country == 'USA') 
  
```

#Network Analysis of Linked Movies

On imdb.com, for each movie, there are 12 recommended movies.  For each movie in this imdb data set, I went to imdb.com and pulled down the 12 recommended movie links.  That's what this below function does.

```{r, eval = T, include = T}

get_links <- function(address) {
  # read the movie page
  page <- readLines(address)
  # find the lines with the recommendations and strip the unneeded stuff
  recs <- page[grep("rec_item", page)]
  recs <- unlist(strsplit(recs, "data-tconst="))[seq(from = 2, to = 24, by = 2)]
  # return the codes
  recs <- paste("tt", gsub("[^0-9]", "", recs), sep = "")
  
  recs = paste("http://www.imdb.com/title/", recs, sep = "")
  return(recs)
}

#Example

get_links(imdb$links[1])

```


Here I take a smalll sample at first to test the system.

```{r, eval = T, include = T}

mydata = filter(imdb, year >= 2016)
#View(mydata2)
dim(mydata)

```

45 movies is a good starting point.  This code below creates an adjacency matrix that is used for the network plot.  The i,jth entry is 1 if movie i is connected to movie j and zero otherwise.

Here is the upper 1-3th quadrant of one such matrix.

```{r}
m = matrix(  c( 0, 1, 0, 1, 0, 1, 0, 1, 0), nrow = 3)

rownames(m) = c("Star Wars", "Toy Story", "Fiddler on the Roof") 
colnames(m) = rownames(m)
m
```


Here Star Wars in connected to Toy Story but not Fiddler on the Roof, and Fiddler on the Roof is connected to Toy Story but not Star Wars.

#Creating a Connections Matrix

```{r, eval = T, include = T}


summary(imdb$genres_simple)
make_network = function(cur_data, cur_size = 10, add.label = T){
    #Due to a restriction on RcolorBrewer, we can only use a max of 9 genres.
    #Here I filter out all but the 9 most popular.  I do so only when necessary
  
    if( length(unique( cur_data$genres_simple)) > 9) {
    cur_data = filter( cur_data, ( genres_simple %in% c("Action",
                                                "Adventure",
                                                "Animation",
                                                "Comedy",
                                                "Crime",
                                                "Drama",
                                                "Fantasy",
                                                "Horror",
                                                "Biography")))
    }
  
    n = nrow(cur_data)
    #Need an index number for each link to match
    vectorize = function(input_links){
      out = c(rep(0, n))
      for(i in input_links){
            index = match(i, cur_data$links)
            out[index] = 1
        }
      return(out)
    }
   
  each_links = sapply(cur_data$links, get_links)
  links_matrix = ldply(each_links, vectorize) %>%
    select(-1) %>%
    as.matrix()
  
   #Add movie titles 
  rownames(links_matrix) = cur_data$title
  colnames(links_matrix) = cur_data$title
  
  #Remove empty connections, ie, movies with no links
  # links_matrix = links_matrix[,which(!apply(links_matrix,2,FUN = function(x){all(x == 0)}))]
  # links_matrix = links_matrix[which(!apply(links_matrix,2,FUN = function(x){all(x == 0)})), ]
  # 
  net1 = network( links_matrix, directed = F, na.rm = F)
  
  # cur_data = filter(cur_data, (title %in% rownames(links_matrix)))
 
  network.vertex.names(net1) = rownames(links_matrix)
  
  length = nrow(links_matrix)
 
  ggnet2(net1, 
         color = cur_data$genres_simple, 
         palette = "Set1",
         label = add.label,
         size = cur_size)
  
  }

```
#Movies in the US released after 2016

You can find interesting patterns between the linked movies over different years.

```{r}
make_network(mydata)
```

#Movies from 1970 to 1980

Here we see a spare connection matrix for the 1970s.  An interesting question is if each movie is related to 12 others, then what are these movies related to?

```{r}
mydata2 = filter(imdb, year > 1970, year < 1980,  country == "USA")
dim(mydata2)

make_network(mydata2)
```

#Movies from Prior to 1975

From this graph we see there are far fewer connections, and it appears movies that are released at the same time are more likely to be related, as are sequels.

```{r}
mydata3 = filter(imdb, year > 1900, year < 1975,  country == "USA")
dim(mydata3)

make_network(mydata3, cur_size = 9)
```

#Movies from 2010 - 2014

Here there are too many movies to be able to read each title, so I just remove it. 

Does it look like movies are generally grouped by genre?

```{r}

mydata4 = filter(imdb, year > 2005, year < 2008,  country == "USA")
dim(mydata4)

make_network(mydata4, cur_size = 3, add.label = F)
```


